---
title: "DRIGRASS -Summary stats for report-"
author: "Shun Hasegawa"
date: "18 December 2016"
output:
  word_document:
    fig_caption: yes
    reference_docx: summary_stats_report_style.docx
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
library(car)
library(lmerTest)
library(ggplot2)
library(grid)
library(pander)
library(plyr)
library(dplyr)
library(tidyr)
library(vegan)
library(lattice)
require(gridExtra)

# knitr::opts_chunk$set(echo = TRUE, fig.path='Figs_report/DRIGRASS_harvest_',
#                       dev=c('pdf', 'png'), dpi = 600)
knitr::opts_chunk$set(echo = TRUE, dpi=600)

load("../Output/Data/all_obj.RData")
load("../Data/all_obj_traits.RData") # results from plant traits

theme_set(theme_bw()) 

# table setting
set.alignment('left', row.names = 'right', permanent = TRUE)

# function to generate anova table from lmer
lmer_anova_tbl <- function(lmermod, ...){
  a <- Anova(lmermod, test.statistic = "F")
  panderOptions('round', c(2, 0, 0, 3))
  if(any(a[, 4] <= 0.05)) emphasize.strong.cells(cbind(which(a[, 4] <= 0.05), 4))
  emphasize.italics.cols(4)
  pander(a, ...)
}

```


## To do
* ~~explain results of forth-corner analysis~~
* mention plant functional groups
* ~~add box-whisker plots of plant traits for each species~~
* ~~replace figures with the ones with better resolution (dpi = 600)~~
* ~~mention species weight on PRC~~
* ~~merge summer and winter PRC plots~~
* ~~merge summer and winter MDS plots~~
* ~~add dominant species~~
* ~~add star plots~~
* ~~mention herbivore effects~~
* (change error bars to ellipses on MDS ordination plots)


<br>

## Summary results
Overall, we found substantial effects of rainfall treatments on plant biomass and community at the study site, whereas there was little evidence of herbivore effects. The impacts of rainfall treatments were most distinctive in the summer in the last year (2016) of this study (Table 1, Table 2).

<br>

```{r smmry_stat_tbl, echo=FALSE, results='asis'}
sst <- ldply(all_ow_anova, function(x) x$summary_stats) %>%
  gather(term, pval, herb, treatment, `treatment:herb`) %>% 
  mutate(pval = ifelse(pval <= .05, as.character(get_star(pval, dagger = FALSE)), "ns"),
         term = recode_factor(term, treatment = "Rain", herb = "Herb", `treatment:herb` = "RxH"),
         SYT  = paste(year, term, sep = "\n")) %>%
  select(-year, -term) %>% 
  spread(SYT, pval) %>% 
  mutate(variable = dplyr::recode_factor(variable, 
                                         live                 = "Live biomass", 
                                         Dead                 = "Dead biomass", 
                                         total                = "Total biomass", 
                                         c34ratios            = "C~3~:C~4~ ratio", 
                                         grprop               = "Grass proportion", 
                                         H                    = "*H'*", 
                                         J                    = "*J'*", 
                                         S                    = "*S*", 
                                         Forks                = "Root forks", 
                                         Tips                 = "Root tips", 
                                         sr_ratio             = "Shoot:root ratios", 
                                         total_L              = "Root length", 
                                         total_SA             = "Root surface area",
                                         Axonopus.fissifolius = "*Axonopus fissifolius*",
                                         Cymbopogon.refractus = "*Cymbopogon refractus*",
                                         Cynodon.dactlyon     = "*Cynodon dactlyon*    ",
                                         Digitaria.sp         = "*Digitaria sp*        ",
                                         Eragrostis.curvula   = "*Eragrostis curvula*  ",
                                         Hypochaeris.radicata = "*Hypochaeris radicata*",
                                         Microlaena.stipoides = "*Microlaena stipoides*",
                                         Paspalum.dilitatum   = "*Paspalum dilitatum*",
                                         Setaria.parviflora   = "*Setaria parviflora*")) %>% 
  arrange(variable)

sst_summer <- sst %>% 
  filter(season == "Summer") %>% 
  select(-season) %>% 
  .[!apply(., 2, function(x) all(is.na(x)))]

sst_winter <- sst %>% 
  filter(season == "Winter") %>% 
  select(-season) %>% 
  .[!apply(., 2, function(x) all(is.na(x)))]
sst_winter[is.na(sst_winter)] <- "-"

pander(sst_summer, split.cell = 2, split.table = Inf, caption = "Table 1. Summary results of analysis of variances (ANVOAs) with herbivore (Herb), rainfall (Rain) and their interaction (RxH) as fixed factors for summer harvests on the following response variables: live, dead, and total biomass, C~3~:C~4~ ratios, the proportion of grasses in the total abundance, diversity indices (Shannon-Wiener diversity index (*H'*), Pielou’s evenness (*J'*) and species richness (*S*)), the community-weighted averages of plant traits  (the number of root forks and tips, total root length and surface area, and shoot:root ratios) and dominant species. The plant traits, except shoot:root ratios, were expresses on root mass basis. Significance codes are *P* < 0.05 (\\*), < 0.01 (\\**), and < 0.001 (\\***), respectively, otherwise shown as ns (not significant). *F* values and associated degrees of freedom used for the ANOVAs are shown in Tables S1.")
```

<br>

```{r Tabl2_stat_winter, echo=FALSE, results='asis'}
pander(sst_winter, split.cell = 2, split.table = Inf, caption = "Table 2. Summary results of analysis of variances (ANVOAs) with herbivore (Herb), rainfall (Rain) and their interaction (RxH) as fixed factors for winter harvests on the following response variables: live, dead, and total biomass, C~3~:C~4~ ratios, the proportion of grasses in the total abundance, diversity indices (Shannon-Wiener diversity index (*H'*), Pielou’s evenness (*J'*) and species richness (*S*)), the community-weighted averages of plant traits  (the number of root forks and tips, total root length and surface area, and shoot:root ratios) and dominant species. The plant traits, except shoot:root ratios, were expresses on root mass basis. Significance codes are *P* < 0.05 (\\*), < 0.01 (\\**), and < 0.001 (\\***), respectively, otherwise shown as ns (not significant). *F* values and associated degrees of freedom used for the ANOVAs are shown in Tables S1.")

```

<br>

## Biomass
### Total, live and dead
ANOVAs demonstrated significant rainfall treatment effects on above-ground live biomass in summer (Table 1, Fig 1), but no effects of herbivore addition. In general, live biomass in SD was significantly lower than the other rainfall treatments, with  72% (*P* < 0.05), 16% (*P* > 0.05) and 58% (*P* < 0.05) decreases relative to AMB treatment in 2014, 2015 and 2016, respectively. The highest live biomass was observed in IR with 23%, 54% and 33% increases relative to AMB in 2014, 2015 and 2016, respectively, although these differences were not significant. There was no evidence of rainfall effects on live biomass in winter (Table 2, Fig 1). Although an ANOVA demonstrated significant effects of herbivore addition in the winter 2013, this was owing to initial spatial heterogeneity at the study site as herbivores had not been applied at that time yet. Total biomass generally followed the patterns observed in live biomass. There was little evidence of rainfall treatment effects on dead biomass. Nevertheless, significant treatment effects were found in the winter 2014 with -32% in RR and -18% in SD compared to AMB, although these differences were not significant.  

<br>

```{r Fig1_biomass, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide', fig.cap="Fig 1. Biomass (Mean ± 95%CI). Confidence intervals (CIs) were estimated from linear models. When response variables were transformed for the linear model analysis (ANOVA), estimated CIs were reverse transformed prior to plotting. Statistically significant differences between rainfall treatments were shown by different characters above associated error bars."}
grid.draw(ab_biom_plot_merged)
```

<br>

### Biomass allocation
We assessed biomass allocation from above- and below-ground comprehensive harvest conducted in September 2015. ANOVAs demonstrated no significant rainfall effects on any of above-, below-ground biomass or above:blow-ground biomass ratios (Fig 2).

```{r Fig2_biomass_allocation, echo=FALSE, warning=FALSE, fig.width=3, fig.height=2, fig.cap="Fig 2. Above- and below-ground biomass (Mean ± 95%CI). Confidence intervals (CIs) were estimated from linear models. When response variables were transformed for the linear model analysis (ANOVA), estimated CIs were reverse transformed prior to plotting."}
fig_tot_biom
```

<br>

## Diversity indices
In summer, significant rainfall treatment effects on *H'* were found in 2014 and 2016 but not in 2015 (Table 1, Fig 3). The highest *H'* was observed in AMB throughout the years. Compared to this, that in RF was lower by 26% (*P* < 0.05), 15% (*P* > 0.05) and 31% (*P* < 0.05) in 2014, 2015 and 2016, respectively. *H'* in the other rainfall treatments was not significantly different than that in AMB. *J'* in summer showed similar trends to *H'*, with significant rainfall effects observed in 2016. Whilst *J'* was significantly decreased in RF by 25%, that in the other rainfall treatments were not significant altered compared to AMB. ANOVAs demonstrated significant rainfall effects on *S* in the summer 2015 and 2016 but not in 2014, with the highest value observed in AMB throughout the years. RR in 2015 and SD in 2016 significantly decreased *S* by 20% and 40% relative to AMB, respectively.

In winter, ANOVAs demonstrated significant rainfall effects on *H'* in 2014 but not in 2013. In 2014, the highest *H'* was observed in SD with 16% increases (*P* > 0.05), whereas the lowest was in RR with 25% decreases (*P* < 0.05) relative to AMB. No evidence of rainfall treatment effects on *J'* was found in winter. *S* showed similar trends to *H'* with significant rainfall treatment effects observed in 2014. RR significantly decreased *S* by 24% compared to AMB. The highest *S* was observed in SD throughout the years with 16% increases relative to AMB, although this difference was not significant.

<br>

```{r Fig3_diversity, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide', fig.cap="Fig 3. Diversity indices (Mean±95%CI). Confidence intervals (CIs) were estimated from linear models. When response variables were transformed for the linear model analysis (ANOVA), estimated CIs were reverse transformed prior to plotting. Statistically significant differences between rainfall treatments were shown by different characters above associated error bars."}
grid.draw(div_plot_merged)
```

<br>

## Dominant species
In summer, dominant species were generally more abundant in AMB and IR compared to RR and SD, with no herbivore effect (Fig 4). For instance, SD significantly increased the abundance of *C. refractus* in 2016 by 92% relative to AMB. ANOVAs demonstrated significant rainfall effects on *M. stipoides* (Table 1) with significant decreases by 98% and 95% under SD in 2014 and 2016 and by 66% under RR in 2015 compared to AMB, respectively. The abundance of *M. stipoides* was highest under IR through the study years with 98% increases relative to AMB, although this difference was not significant. An ANOVA also demonstrated significant rainfall effects on *P. dilitatum* in 2014 with significant decreases by 81% under SD compared to AMB. In winter, on the other hand, there was no evidence of rainfall effects on any of the dominant species (Fig 4, Table 2). An ANOVA demonstrated significant effects of herbivore addition on the abundance of *H. radicata* in 2014 with increases by 93% compared to control treatment.

<br>

```{r Fig4_dominant, echo=FALSE, fig.height=5, fig.width=6.5, fig.cap="Fig 4. Star plots of dominant species (*Axonopus fissifolius*, *Cymbopogon refractus*, *Cynodon dactlyon*, *Digitaria sp*, *Eragrostis curvula*, *Hypochaeris radicata*, *Microlaena stipoides*, *Paspalum dilitatum* and *Setaria parviflora*) for each rainfall treatment: Ambient (AMB), Increased (IR), Reduced (RR), Reduced frequency (RF) and Summer drought (SD)."}
plot_starplot()
```

<br>

## Plant functional groups
```{r Fig_pfg, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide'}
grid.draw(pfg_fig_merged)
```

<br>

## MDS analysis
We performed MDS to visualise the whole patterns of the plant communities for summer and winter (Fig 5). The first two MDS axes (MDS1 and MDS2) explained 29% and 11% of the total variation for summer and 28% and 16% for winter, respectively. For summer, MDS1 was mainly associated with annual changes in the community, whereas MDS2 was associated with rainfall treatments. The separations between rainfall treatments grew larger with study years, indicating that rainfall treatments caused successions in the community. Consequently, IR and SD distinctively deviated from the other rainfall treatment in 2015 and 2016, respectively. For winter, a large proportion of the variation in the community was associated with annual changes. With respect to rainfall treatments, whilst no clear separation was demonstrated in the MDS ordination in 2013, SD diverged from the others in 2014. 

```{r Fig5_mds, echo=FALSE, fig.height=5, fig.width=6.5, results='hide', fig.cap="Fig 5. Ordination plot of multidimensional scaling (MDS) on the plant communities in summer and winter. The proportions of variation of the community explained by the first two axes of MDS (MDS1 and MDS2) are shown in parentheses."}
grid.draw(merged_MDS)
```

<br>

## Principal response curve (PRC) and PERMANOVA
We performed principal response curve (PRC) models on the first principal axis for summer and winter (Fig 7). Permutation tests demonstrated significant time-dependent rainfall treatment effects on the communities for both of summer (*F*(1, 75)=5.0, *P* < 0.05) and winter (*F*(1, 50)=3.5, *P* < 0.05). Furthermore, we performed PERMANOVAs for each harvesting event separately. This demonstrated significant rainfall effects on the community in the summer 2016 (Table 3, *F*(4,25) = 1.9, *P* < 0.05), but not for the other harvests. The PRC diagrams for summer and winter generally supported the trends demonstrated in MDS ordination plots along the MDS1 (Fig 5) and in the analyses for dominant species (Fig 4). The community under SD increasingly diverged from the other rainfall treatments with the study years for both of summer and winter. In summer, this was especially derived from increased abundance of *C. dactlyon* decreased abundance of *M. stipoides* under SD relative to AMB (Fig 4, Fig 7).

<br>

```{r Fig7, echo=FALSE, fig.height=4, fig.width=6.5, results='asis', fig.cap="Fig 7. Principal response curves (PRCs) for summer and winter harvests. Species weight are shown along the right y axis for species with the total abundance >100 g m^-2^: *Axonopus fissifolius*, *Bothriochloa macra*, *Cymbopogon refractus*, *Cynodon dactlyon*, *Digitaria sp.*, *Eragrostis curvula*, *Lolium perenne*, *Microlaena stipoides*, *Paspalum dilitatum*, *Plantago lanceolata*, *Setaria parviflora* and *Sida rhombifolia*"}
plot_prc_fig()
```


```{r permanova, echo=FALSE, results='asis'}
p_tbl <- prmnv_res_tbl %>% 
  mutate(Year     = gsub("_.*", "", .id),
         Season   = gsub(".*_", "", .id),
         pseudo_F = gsub("F", "*F*", pseudo_F),
         R2       = round(R2, 2),
         P        = round(P, 3)) %>% 
  rename(`*R^2^*`     = R2,
         `Psuedo-*F*` = pseudo_F,
         `*P*`        = P) %>%
  mutate(Season = ifelse(duplicated(Season), "", Season)) %>% 
  select(Season, Year, `Psuedo-*F*`, everything(), -.id)
emphasize.italics.cols(5)
emphasize.strong.cells(cbind(which(p_tbl[,5] < .05), 5))
pander(p_tbl, caption = "Table 3. Results of PERMANOVA for each harvest for each year with 9999 permutations. The Bray-Curtis dissimilarity was used.")
```

<br>

## Plant traits
We assess the effects of rainfall treatments on plant traits: specific number of root forks and tips, root length and surface area and shoot:root ratios. There was little evidence of rainfall effects on the traits within the same species. Nevertheless, significant rainfall effects on specific root length and the number of tips and forks were found for *Axonopus fissifolius* but not for the other species. All the assessed plant traits substantially varied between species (Fig 9, Table S2, *P* < 0.001). In general, *Lolium perenne* demonstrated high values for the root traits, indicating that it had thinner roots compared to the other species. Although *Bromus* and *Lotus* also demonstrated relatively high values for the root traits, there was only single measurement; hence, within-species variations were not captured. By contrast, *Hypochaeris radicata*, *Microlaena stipoides* and *Paspalum dilatatum* showed relatively low values for the root traits. indicating thick roots. With respect to shoot:root ratios, *Eragrostis* and *Vicia* demonstrated high ratios, whilst *Hypochaeris radicata* and *Microlaena stipoides* low ratios. We further assessed the community-level plant traits by computing abundance-weighted averages for each harvesting event. There was little evidence of rainfall treatment effects on any of the assessed plant traits at the community level (Table 1, Table 2, Fig 8). Nevertheless, ANOVA demonstrated significant treatment effects on the specific number of the root forks at the community level in winter 2014 (Table 2).

<br>

```{r Fig8_plant_trait, echo=FALSE, fig.height=8.5, fig.width=6.5, results='asis', fig.cap="Fig 8. Abundance-weighted means of community-level plant traits."}
grid.draw(trait_plot_merged)
```

<br>

```{r Fig9_trait_bySpp, echo=FALSE, fig.height=4, fig.width=6.5, results='asis', fig.cap="Fig 9.  Box-whisker plots of plant traits summarised by species."}
grid.draw(trait_fig_bySpp_merged)
```


<br>

## Fourth corner models (community assembly via trait selection)  
Fourth corner models demonstrated significant interaction effects between rainfall treatment and plant traits on the plant community in the winter 2014, indicating strong association between plant traits and community successions induced by rainfall treatment. Although this was not significant, the final model with the lowest BIC for the summer 2015 kept RR:root length term in the model (Table 8, Fig 10). In the winter 2014, SD was associated positively with specific root length and negatively with shoot:root ratios (Fig 10). This indicated that the species with thin roots and/or low shoot:root ratios (e.g. *L. perenne* and *P .lanceolata*, Fig 9) were relatively more abundant, whereas those with thick roots and/or low shoot:root ratios (e.g. *E. curvula* and *M. stipoides*, Fig 9) were less in SD compared to AMB. In the summer 2015, specific root length was positively associated with RR compared to AMB. This indicated that those thin roots (e.g. *C dactlyon* and *S. parviflora*, Fig 9) were relatively more abundant, whereas those with thick roots (e.g. *M stipoides* and *P. dilitatum*, Fig 9) were less in RR compared to AMB. The results of forth corner models generally support the trend demonstrated in the PRC models (Fig 7).


<br>

```{r Table8_4thcorner, echo=FALSE, results='asis'}

rql_res_tbl <- rql_res %>% 
  mutate(Season = ifelse(duplicated(Season), "", Season),
         Dev = round(Dev, 2)) %>%
  rename(`*Pr*(>Dev)` = `Pr(>Dev)`)
rql_res_tbl[is.na(rql_res_tbl)] <- "-"
  
emphasize.strong.cells(cbind(which(rql_res_tbl[, 6] <= 0.05), 6))
emphasize.italics.cols(6)
pander(rql_res_tbl, caption = "Table 8. Summary results of environment-trait terms (i.e. interaction between rainfall treatment and plant traits) in fourth corner models. No interaction terms were kept in the final model with the lowest BIC for the summer 2014 and 2016 and the winter 2013.")
```


```{r Fig10_4thcorner, echo=FALSE, fig.width=6, fig.height=5, fig.cap="Fig 10. Standardised interaction coefficient estimates for interaction terms between rainfall treatment and plant traits from fourth corner analysis with significant environment-trait interactions (see Table 8). The associations for interaction terms in the best model with the lowest BIC are shown in blue or red with color strength representing the strength of associations. Red and blue represent positive and negative association in comparison to control treatment (AMB), respectively. Rainfall treatments are Increased (IR), Reduced (RR), Reduced frequency (FR) and Summer drought (SD). Root length and tips are root mass basis. No interaction terms were kept in the final model for the summer 2014 and 2016 and the winter 2013."}
grid.arrange(rql_plt_l[[1]],rql_plt_l[[2]], ncol=1)
```



## Supplementary
```{r TableS1_anova_tbl, echo=FALSE, results='asis', warning=FALSE}
summary_tbl_ed <- summary_tbl %>%
  mutate(variable = dplyr::recode_factor(variable, 
                                         live                 = "Live biomass", 
                                         Dead                 = "Dead biomass", 
                                         total                = "Total biomass", 
                                         c34ratios            = "C~3~:C~4~ ratio", 
                                         grprop               = "Grass proportion", 
                                         H                    = "*H'*", 
                                         J                    = "*J'*", 
                                         S                    = "*S*", 
                                         Forks                = "Root forks", 
                                         Tips                 = "Root tips", 
                                         sr_ratio             = "Shoot:root ratios", 
                                         total_L              ="Root length", 
                                         total_SA             = "Root surface area",
                                         Axonopus.fissifolius = "*Axonopus fissifolius*",
                                         Cymbopogon.refractus = "*Cymbopogon refractus*",
                                         Cynodon.dactlyon     = "*Cynodon dactlyon*",
                                         Digitaria.sp         = "*Digitaria sp*",
                                         Eragrostis.curvula   = "*Eragrostis curvula*",
                                         Hypochaeris.radicata = "*Hypochaeris radicata*",
                                         Microlaena.stipoides = "*Microlaena stipoides*",
                                         Paspalum.dilitatum   = "*Paspalum dilitatum*",
                                         Setaria.parviflora   = "*Setaria parviflora*"),
         `F` = gsub("F", "*F*", `F`)) %>% 
  arrange(variable) %>%
  rename(`*F*-test` = `F`,
         `*P*`      = P,
         Year       = year) %>% 
  mutate(Variable = ifelse(duplicated(variable), "", as.character(variable))) %>% 
  group_by(variable) %>% 
  mutate(Season = ifelse(duplicated(season), "", as.character(season))) %>% 
  ungroup() %>% 
  select(Variable, Season, Year, everything(), -variable, -season) 

emphasize.strong.cells(cbind(which(summary_tbl_ed[, 5] <= 0.05), 5))
emphasize.italics.cols(5)
pander(summary_tbl_ed[, 1:5], split.table = Inf, caption = "Table S1. Results of one-way ANOVAs. Degrees of freedom used for *F* tests are shown in parentheses")

```

<br>

```{r TableS2_plant_traits, echo=FALSE, warning=FALSE}
a <- smmry_anova_bySpp_tbl %>%
  mutate(`F` = gsub("F", "*F*", `F`), 
         unit = gsub("mg-1", "mg^-1^", unit), 
         trait = dplyr::recode_factor(trait, 
                               length = "Root length",
                               `surface area` = "Root SA",
                               forks = "Root forks",
                               tips = "Root tips",
                               shoot_root_ratio = "Shoot:root ratio"),
         P = "_**<0.001**_") %>% 
  gather(spp, value, -trait, -unit, -F, -P) %>% 
  gather(variable, value2, -spp, -trait) %>% 
  spread(trait, value2) %>% 
  filter(!(spp != "Axonopus" & variable %in% c("F", "P", "unit"))) %>% 
  mutate(variable = ifelse(variable == "value", spp, variable),
         variable = dplyr::recode(variable,
                                  F = "*F*-test",
                                  P = "*P*")) %>% 
  rename(" " = variable) %>% 
  select(-spp)

pander(a, split.table = Inf, caption = "Table S2. Results of one-way ANOVA for species comparisons and summary with Mean±SE. SE is NA when *n* < 2. Where there was a significant treatment effect at *P* < 0.05, post-hoc tests were performed. The results of post-hoc comparisons are shown by letters alongside parentheses")
```
