---
title: "DRIGRASS -Summary stats -"
author: "Shun Hasegawa"
date: "17 October 2016"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(car)
library(lmerTest)
library(ggplot2)
library(grid)
library(pander)
library(plyr)
library(dplyr)
library(vegan)
knitr::opts_chunk$set(echo = TRUE, fig.path='Figs/DRIGRASS_harvest_',
                      dev=c('pdf', 'png'), dpi = 600)
load("../Output/Data/all_obj.RData")
theme_set(theme_bw()) 


# function to generate anova table from lmer
lmer_anova_tbl <- function(lmermod, ...){
  a <- Anova(lmermod, test.statistic = "F")
  panderOptions('round', c(2, 0, 0, 3))
  if(any(a[, 4] <= 0.05)) emphasize.strong.cells(cbind(which(a[, 4] <= 0.05), 4))
  emphasize.italics.cols(4)
  pander(a,  justify = "right", ...)
}

```



## Introduction
Here, I show the summary results of statistical analysis for the DRIGRASS experiment. Repeated-measures ANOVA was performed separately for summer (Apr) and winter (Oct) data. Firstly, interaction effects between rainfall and herbivore were assessed using a subset of the data with the rainfall treatments of Pulsed.drought, Ambient and Drought. Secondly, where there was a significant Herbivore effect (indicated as either main or interaction effect), I removed added-herbivore plots from the dataset and performed repeated-measures ANOVA for rainfall treatments (_n_ = 6). Where there was no Herbivore effect, the full dataset was analysed for rainfall treatment (_n_ = 6 or 12). Also, I perfomed one-way ANOVA for each season in each year individually. 




## Biomass (Total/Live/Dead)

```{r biom_fig, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide'}
grid.draw(ab_biom_plot_merged)
```

### Total biomass


#### Summer
```{r biom_tbl_summer, echo=FALSE, results='asis', fig.width=5, fig.height=6, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(total_rh_s_m_fin, 
           caption = "Total biomass in summer for rainfall and herbivore treatments")

lmer_anova_tbl(total_s_m_fin, 
           caption = "Total biomass in summer for rainfall treatments")
plot(lsmeans::lsmeans(total_s_m_fin, by = "year", specs = "treatment"), comparisons = TRUE)
```

#### Winter
```{r biom_tbl_winter, echo=FALSE, results='asis', fig.width=3, fig.height=2.5, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(total_rh_w_m_fin, 
           caption = "Total biomass in winter for rainfall and herbivore treatments")

lmer_anova_tbl(total_w_m_fin, 
           caption = "Total biomass in winter for rainfall treatments")
plot(lsmeans::lsmeans(total_w_m_fin, "year"), comparisons = TRUE)
```




## Diversity indices

```{r div_fig, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide'}
grid.draw(div_plot_merged)
```


### Shannon-weaver index (_H'_)


#### Summer

```{r h_s_tbl_rain, echo=FALSE, results='asis', fig.width=3, fig.height=2.5, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(h_by_rh_s_m_fin, caption = "Shannon-weaver index in summer for rainfall and herbivore treatments")

lmer_anova_tbl(h_s_m_fin, caption = "Shannon-weaver index in summer for rainfall")
plot(lsmeans::lsmeans(h_s_m_fin, "treatment"), comparisons = TRUE)
```




#### Winter
```{r h_w_tbl, echo=FALSE, results='asis', fig.width=5, fig.height=4, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(h_by_rh_w_m_fin, caption = "Shannon-weaver index in winter for rainfall and herbivore treatments")

lmer_anova_tbl(h_w_m_fin, caption = "Shannon-weaver index in winter for rainfall")
plot(lsmeans::lsmeans(h_w_m_fin, by = "year", specs = "treatment"), comparisons = TRUE)

```




### Evenness (_J'_)

#### Summer
```{r j_s_tbl_rxh, echo=FALSE, results='asis', fig.width=5, fig.height=5, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(j_by_rh_s_m_fin, caption = "Species evennes in summer for rainfall and herbivore treatments")
plot(lsmeans::lsmeans(j_by_rh_s_m_fin, by = "treatment", specs = "herb"), comparisons = TRUE)
```


```{r j_s_tbl_rain, echo=FALSE, results='asis'}
j_s_m_fin@call
```
None of fixed term was significant.


#### Winter
```{r j_w_tbl, echo=FALSE, results='asis', fig.width=3, fig.height=2.5, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(j_by_rh_w_m_fin, caption = "Species evennes in winter for rainfall and herbivore treatments")
lmer_anova_tbl(j_w_m_fin, caption = "Species evennes in winter for rainfall")
plot(lsmeans::lsmeans(j_w_m_fin, "year"), comparisons = TRUE)
```



### Species richness


#### Summer
```{r s_s_tbl, echo=FALSE, results='asis', fig.width=3, fig.height=2.5, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(s_by_rh_s_m_fin, caption = "Species richness in summer for rainfall and herbivore treatments")

lmer_anova_tbl(s_s_m_fin, caption = "Species richness in summer for rainfall")
plot(lsmeans::lsmeans(s_s_m_fin, "treatment"), comparisons = TRUE)
plot(lsmeans::lsmeans(s_s_m_fin, "year"), comparisons = TRUE)
```


#### Winter
```{r s_w_tbl, echo=FALSE, results='asis', fig.width=3, fig.height=2.5, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(s_by_rh_w_m_fin, caption = "Species richness in winter for rainfall and herbivore treatments")

lmer_anova_tbl(s_w_m_fin, caption = "Species richness in winter for rainfall")
plot(lsmeans::lsmeans(s_w_m_fin, "treatment"), comparisons = TRUE)
plot(lsmeans::lsmeans(s_w_m_fin, "year"), comparisons = TRUE)
```


## Plant functional groups


```{r pfg_fig, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide'}
grid.draw(pfg_fig_merged)
```

### C3:C4 ratios


#### Summer
```{r c34r_s_tbl, echo=FALSE, results='asis', fig.width=5, fig.height=6, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(c34r_by_rh_s_fin, caption = "C3:C4 ratios in summer for rainfall and herbivore treatments")

lmer_anova_tbl(c34r_s_m_fin, caption = "C3:C4 ratios in summer for rainfall treatments")
plot(lsmeans::lsmeans(c34r_s_m_fin, by = "year",spec = "treatment"), comparisons = TRUE)
```


#### Winter
```{r c34r_w_tbl, echo=FALSE, results='asis', fig.width=5, fig.height=6, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(c34r_by_rh_w_fin, caption = "C3:C4 ratios in winter for rainfall and herbivore treatments")

lmer_anova_tbl(c34r_w_m_fin, caption = "C3:C4 ratios in winter for rainfall treatments")
plot(lsmeans::lsmeans(c34r_w_m_fin, by = "year",spec = "treatment"), comparisons = TRUE)
```



### Grass proportion

#### Summer
```{r grprop_s_tbl, echo=FALSE, results='asis', fig.width=5, fig.height=6, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(grprop_by_rh_s_fin, caption = "Grass proportion in summer for rainfall and herbivore treatments")

lmer_anova_tbl(grprop_s_m_fin, caption = "Grass proportion in summer for rainfall treatments")
```


#### Winter
```{r grprop_w_tbl, echo=FALSE, results='asis', fig.width=5, fig.height=6, fig.cap="Least square mean comparisons. Estimated mean ± 95%CI. There is a significant difference when arrows are not overlapped.", message=FALSE, fig.show='hold'}
lmer_anova_tbl(grprop_by_rh_w_fin_2, caption = "Grass proportion in winter for rainfall and herbivore treatments")

lmer_anova_tbl(grprop_w_m_fin, caption = "Grass proportion in winter for rainfall treatments")
plot(lsmeans::lsmeans(grprop_w_m_fin, by = "year",spec = "treatment"), comparisons = TRUE)

```


## One-way ANOVA for each harvest in each year
```{r oneway_anova_tbl, echo=FALSE, results='asis', warning=FALSE}
emphasize.strong.cells(cbind(which(summary_tbl[, 5] <= 0.05), 5))
emphasize.italics.cols(5)
pander(summary_tbl, justify = "right", split.table = Inf, caption = "Results of one-way ANOVA and summary with mean±SE (n = 6 [control-herb] or 12[added- & control-herb]). Where there was a significant treatment effect at P < 0.05, post-hoc tests were performed. The results of post-hoc comparisons are shown by letters alongside parenthesis.")

```


## Principal response curve

```{r prc, echo=TRUE, results='markup', warning=FALSE, fig.height=4, fig.width=6.5}
summer_df <- ab_spp_biom %>% 
  filter(treatment != "No.shelter" & herb == "Control" & month == 4) %>% 
  arrange(plot, year)

spssum <- colSums(summer_df[spp_names])
sp_summ_df <- summer_df[, spp_names[spssum != 0]]

yearv <- factor(summer_df$year)
treatv <- factor(summer_df$treatment)


pc_df_s <- decostand(sp_summ_df, method = "hellinger")
prc_smmr <- prc(pc_df_s, treatment = treatv, time = yearv) 
plot(prc_smmr)
contr <- how(within = Within(type = "series"),
             plots  = Plots(strata = summer_df$plot, type = "free"),
             nperm  = 999)
anova(prc_smmr, permutations = contr)


```


<!-- ## Dominent species (>5% of total biomass) -->

<!-- ```{r dom_spp_fig, echo=FALSE, fig.height=7, fig.width=7, results='hide'} -->
<!-- sp_plot -->
<!-- ``` -->

<!-- ```{r dom_spp_tbl, echo=FALSE, results='asis'} -->
<!-- l <- llply(names(m_rain_list), function(x) pander(anova(m_rain_list[[x]]), justify = "right", caption = x)) -->
<!-- l[[1]] -->
<!-- l[[2]] -->
<!-- l[[3]] -->
<!-- l[[4]] -->
<!-- ``` -->

<!-- ```{r starplot, echo=FALSE, fig.width=8, fig.height=6, results='hide'} -->
<!-- plot_starplot() -->
<!-- ``` -->



<!-- ## NMDS and permANOVA -->

<!-- ```{r nmds_fig, echo=FALSE, fig.height=6, fig.width=7, results='hide'} -->
<!-- plot_nmds_2016() -->
<!-- ``` -->


<!-- ```{r perm, echo=FALSE, results='asis'} -->
<!-- pander(perm_rxh$aov.tab, justify = "right", caption = "permANOVA for rainfall x herb interaction") -->
<!-- pander(perm_rain$aov.tab, justify = "right", caption = "permANOVA for rainfall treatment") -->
<!-- ``` -->


